---
title: "2017 vs 2020 analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(jsonlite)
library(ineq)
library(scales)
library(vroom)
library(lubridate)

#create function to calculate modal value
mode_func <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

```


# Initial wrangling
## Read in data
```{r}
# read in the company information
company_info <- fromJSON("data/company_data_list_9_4_2018.json") %>%
  as_tibble() %>%
  rename(company = owner_name) %>%
  select(company, country, root_parent) %>%
  mutate(country = str_to_upper(country)) %>%
  mutate(leaf_parent = ifelse(is.na(root_parent) | root_parent == "", company, root_parent))

# read in our mapping of genres to super genres
genre_grouping <- read_csv("data/genre_grouping.csv") %>% select(-numApps)

# a file containing a mapping between 2017 and 2020 ids
mapping <- vroom("data/mapping.csv")

apps_2017 <- vroom("data/2017_study/appInfo.csv", 
                   col_types = list(col_integer(), 
                                    col_character(), 
                                    col_character())) %>% 
  mutate(crawl = "2017")

apps_2020 <- vroom("data/xray/apps_all.csv") %>%
  select(-title, -version) %>% 
  mutate(crawl = "2020")

apps <- apps_2017 %>% 
  bind_rows(apps_2020)

hosts_2017 <- vroom("data_processed/2017_hosts_and_companies_long.csv") %>% 
  mutate(crawl = "2017")

hosts_2020 <- vroom("data_processed/2020_hosts_and_companies_long.csv") %>% 
  mutate(crawl = "2020")

hosts <- hosts_2017 %>% 
  bind_rows(hosts_2020)

```

# Main analysis
## Summary statistics of hosts / domains that are associated with tracker companies
```{r}
# count tracker hosts
count_tracker_hosts <- hosts %>% 
  filter(company != "unknown") %>% 
  group_by(crawl, id) %>% 
  summarise(num_tracker_hosts = n())

# set count to 0 for apps that don't have any domains associated with tracker companies
apps_w_no_known_trackers <- apps %>%
  select(crawl, id) %>% 
  anti_join(count_tracker_hosts) %>% 
  mutate(num_tracker_hosts = 0)

# put together in common data frame
count_tracker_hosts <- count_tracker_hosts %>% 
  bind_rows(apps_w_no_known_trackers)

# calculate summary statistics
count_tracker_hosts %>%
  group_by(crawl) %>% 
  summarise(num_apps = n(),
            median = median(num_tracker_hosts),
            Q1 = quantile(num_tracker_hosts, .25),
            Q3 = quantile(num_tracker_hosts, .75),
            IQR = IQR(num_tracker_hosts),
            min = min(num_tracker_hosts),
            max = max(num_tracker_hosts),
            mode = mode_func(num_tracker_hosts),
            mean = round(mean(num_tracker_hosts),1),
            SD = round(sd(num_tracker_hosts),2),
            num_more_than_20 = sum(num_tracker_hosts > 20),
            pct_more_than_20 = round((num_more_than_20 / num_apps) * 100,2),
            no_refs = sum(num_tracker_hosts == 0),
            pct_none = round((no_refs / num_apps) * 100,2)) %>%
  select(-num_more_than_20, -no_refs)

```

Great, now let's also get lorenz curve gini coefficient

```{r}
count_tracker_hosts %>% 
  group_by(crawl) %>% 
  summarise("Gini coefficient" = ineq(num_tracker_hosts, type = 'Gini'))

#plot the distribution in a histogram
count_tracker_hosts %>%
  #filter(num_tracker_hosts < 65) %>%
  ggplot(aes(num_tracker_hosts, fill = crawl)) +
  #geom_histogram(bins = 65) +
  geom_density(alpha = 0.6) +
  labs(x = "Number of tracker hosts per app", 
       title = "Number of tracker hosts in 2017 vs 2020 crawl (log scale)") +
  scale_x_log10()

```


## Most frequent hosts
### tracker hosts on our list
```{r}
#create short mapping from hosts to companies
hosts_to_company <- hosts %>%
  select(-id) %>%
  distinct(hosts, company)

#summary of tracker hosts
known_trackers_info <- hosts %>%
  filter(company != "unknown") %>%
  group_by(hosts, crawl) %>% 
  summarise(num_apps_present = n()) %>% 
  mutate(pct_apps_present = ifelse(crawl == "2017",
                                   (num_apps_present / nrow(apps_2017)*100),
                                   (num_apps_present / nrow(apps_2020)*100))) %>% 
  left_join(hosts_to_company, by = "hosts") %>% 
  left_join(company_info, by = "company") %>%
  arrange(desc(num_apps_present))

known_trackers_info %>% 
  group_by(crawl) %>% 
  arrange(desc(pct_apps_present)) %>% 
  slice(1:30) %>% 
  ungroup() %>%
  ggplot() +
    geom_col(aes(x = reorder(hosts, pct_apps_present), y = pct_apps_present)) +
    facet_wrap(~crawl) +
    coord_flip()

```

### Hosts not on our tracker list
```{r}
#create summary of 'unknown' hosts (i.e. not on our tracker list)
unknown_hosts_info <- hosts %>%
  filter(company == "unknown") %>%
  group_by(hosts, crawl) %>%
  summarise(num_apps_present = n()) %>% 
  mutate(pct_apps_present = ifelse(crawl == "2017",
                                   (num_apps_present / nrow(apps_2017)*100),
                                   (num_apps_present / nrow(apps_2020)*100))) %>% 
  arrange(desc(num_apps_present))

unknown_hosts_info %>% 
  group_by(crawl) %>% 
  arrange(desc(pct_apps_present)) %>% 
  slice(1:30) %>% 
  ungroup() %>% 
  ggplot() +
    geom_col(aes(x = reorder(hosts, pct_apps_present), y = pct_apps_present)) +
    facet_wrap(~crawl) +
    coord_flip()
```


## Distinct companies per app
```{r}
#count number of distinct companies in apps that include hosts that are on our tracker list
company_counts_in_apps_w_known_trackers <- hosts %>%
  filter(company != "unknown") %>%
  group_by(crawl, id) %>%
  distinct(company) %>%
  summarise(num_companies = n())

#set number of companies to 0 for the remaining apps
apps_w_no_known_tracker_hosts <- apps %>%
  select(crawl, id) %>% 
  anti_join(company_counts_in_apps_w_known_trackers) %>% 
  mutate(num_companies = 0)

# join the two
count_tracker_companies <- company_counts_in_apps_w_known_trackers %>% 
  bind_rows(apps_w_no_known_tracker_hosts)

```

Now we can calculate summary statistics

```{r}
#calculate summary statistics of distinct tracker companies per app
count_tracker_companies %>%
  group_by(crawl) %>% 
  summarise(num_apps = n(),
            median = median(num_companies),
            Q1 = quantile(num_companies, .25),
            Q3 = quantile(num_companies, .75),
            mode = mode_func(num_companies),
            min = min(num_companies),
            max = max(num_companies),
            IQR = IQR(num_companies),
            mean_companies = round(mean(num_companies),1),
            SD = round(sd(num_companies),2),
            num_more_than_10 = sum(num_companies > 10),
            pct_more_than_10 = round((num_more_than_10 / num_apps) * 100,2),
            no_refs = sum(num_companies == 0),
            pct_none = round((no_refs / num_apps) * 100,2)) %>%
  select(-num_more_than_10, -no_refs)

```

Explore inequality

```{r}
count_tracker_companies %>% 
  group_by(crawl) %>% 
  summarise("Gini coefficient" = ineq(num_companies, type = 'Gini'))

#plot the distribution in a histogram
count_tracker_companies %>% 
  filter(num_companies < 30) %>%
  ggplot(aes(num_companies, fill = crawl)) +
    #geom_histogram(position = "dodge") +
    geom_boxplot()
    labs(x = "Number of distinct companies per app", y = "Number of apps")
    #scale_x_log10()

count_tracker_companies %>%
  filter(num_companies < 30) %>%
  ggplot(aes(num_companies, fill = crawl)) +
  #geom_histogram(bins = 65) +
  geom_density(alpha = 0.6) +
  labs(x = "Number of tracker hosts per app", 
       title = "Number of tracker hosts in 2017 vs 2020 crawl (log scale)")

```

### Outliers (companies w/ more than 30 )
```{r}
# explore extreme outliers
count_tracker_companies %>%
  filter(num_companies > 30) %>%
  left_join(apps) %>%
  arrange(desc(num_companies))
```


## Presence of specific tracker companies in apps
### Companies
```{r}
#calculate how many percent of apps each company (immediate owner) is present in
prop_apps_w_tracking_company_refs <- hosts %>%
  group_by(crawl, id) %>%
  distinct(company) %>% #exclude the distinct refs within each group
  ungroup() %>% 
  filter(company != "unknown") %>% 
  count(crawl, company) %>% 
  mutate(pct_of_apps = ifelse(crawl == "2017",
                                   (n / nrow(apps_2017)*100),
                                   (n / nrow(apps_2020)*100))) %>% 
  arrange(desc(n))

prop_apps_w_tracking_company_refs %>% 
  group_by(crawl) %>% 
  slice(1:30) %>% 
  ggplot(aes(x = reorder(company, pct_of_apps), y = pct_of_apps)) +
    geom_col() +
    facet_wrap(~crawl) +
    coord_flip() +
    labs(x = "",
         y = "",
         title = "Percentage presence of companies")

```

### Leaf parents
```{r}
#calculate how many percent of apps each root company is present in
prevalence_of_root_companies <- hosts %>%
  filter(company != "unknown") %>%
  left_join(company_info, by = "company") %>%
  distinct(crawl, id, leaf_parent) %>%
  count(crawl, leaf_parent) %>%
  mutate(pct_of_apps = ifelse(crawl == "2017",
                                   (n / nrow(apps_2017)*100),
                                   (n / nrow(apps_2020)*100))) %>% 
  arrange(desc(n))


prevalence_of_root_companies %>% 
  group_by(crawl) %>% 
  slice(1:30) %>% 
  ungroup() %>% 
  ggplot(aes(x = reorder(leaf_parent, pct_of_apps), y = pct_of_apps)) +
    geom_col() +
    facet_wrap(~crawl) +
    coord_flip() +
    labs(x = "",
         y = "",
         title = "Percentage presence of companies")
```


## Analyses by 'super genre'
### NUMBER OF DISTINCT TRACKER COMPANIES PER APP
```{r}
#first describe the number of distinct tracker companies per app for family apps
fam_count_company_refs <- count_tracker_companies %>%
  left_join(apps) %>%
  filter(!is.na(family_genre)) %>%
  mutate(super_genre = "Family")

fam_summary_company_count <- fam_count_company_refs %>%
  group_by(crawl) %>% 
  summarise(num_apps = n(),
            median = median(num_companies),
            Q1 = quantile(num_companies, .25),
            Q3 = quantile(num_companies, .75),
            mode = mode_func(num_companies),
            min = min(num_companies),
            max = max(num_companies),
            IQR = IQR(num_companies),
            mean_companies = round(mean(num_companies),1),
            SD = round(sd(num_companies),2),
            num_more_than_10 = sum(num_companies > 10),
            pct_more_than_10 = round((num_more_than_10 / num_apps) * 100,2),
            no_refs = sum(num_companies == 0),
            pct_none = round((no_refs / num_apps) * 100,2)) %>%
  select(-num_more_than_10, -no_refs) %>%
  mutate(super_genre = "Family")

```

analyse overall

```{r}
#then describe the number of tracker companies per app by super genre, and add a row with the description for family apps to the output
count_tracker_companies %>%
  left_join(apps) %>% 
  left_join(genre_grouping) %>% 
  group_by(crawl, super_genre) %>% 
  summarise(num_apps = n(),
            median = median(num_companies),
            Q1 = quantile(num_companies, .25),
            Q3 = quantile(num_companies, .75),
            mode = mode_func(num_companies),
            min = min(num_companies),
            max = max(num_companies),
            IQR = IQR(num_companies),
            mean_companies = round(mean(num_companies),1),
            SD = round(sd(num_companies),2),
            num_more_than_10 = sum(num_companies > 10),
            pct_more_than_10 = round((num_more_than_10 / num_apps) * 100,2),
            no_refs = sum(num_companies == 0),
            pctNone = round((no_refs / num_apps) * 100,2)) %>% 
  select(-num_more_than_10, -no_refs) %>% 
  bind_rows(fam_summary_company_count) %>%
  arrange(desc(median), desc(Q3), desc(pct_more_than_10), desc(crawl))


```

visualise number of companies

```{r}
#visualise this in a box plot
count_tracker_companies %>%
  left_join(apps) %>%
  left_join(genre_grouping) %>% 
  bind_rows(fam_count_company_refs) %>%
  mutate(super_genre = factor(super_genre,
                              levels = c("Productivity & Tools","Communication & Social","Education","Health & Lifestyle","Music","Art & Photography","Games & Entertainment", "Family","News"), ordered = TRUE)) %>% 
  ggplot(aes(y = num_companies, x = super_genre, fill = crawl)) +
    geom_boxplot(varwidth = TRUE, outlier.shape = NA) + 
    labs(x = "Super genre", y = "Number of distinct tracker companies per app") +
    scale_y_continuous(breaks = seq(0, 20, 4)) +
    coord_flip(ylim = c(0,22)) +
    theme_minimal()

```


# TODO